### 2019-01-16 presentations/January2019/presentation4.tex
asal, nonASAL remove insig
asal, nonASAL remove insig

%\documentclass[notes]{beamer}       % print frame + notes
%\documentclass[notes=only]{beamer}   % only notes
\documentclass{beamer}              % only frames
  \usepackage[english]{babel}
\usepackage[round]{natbib}
%\documentclass[hyperref={pdfpagelabels=false}]{beamer}
\usepackage{lmodern}
\usepackage{marvosym} % \MVRIGHTarrow

\usepackage{amsmath}  
\usepackage{multirow,graphicx,array}
\usepackage{multicol}
\usepackage{hhline}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{threeparttable,booktabs}
\usepackage{chngcntr}
\usepackage{dcolumn}
\usepackage{caption}
\usepackage{fixltx2e}
\usepackage{rotating}
\usepackage{amssymb}

\usetheme{Singapore}
\useoutertheme{miniframes}
\AtBeginSection[]{\subsection{}}

\newcommand{\RowColor}{\rowcolor{gray} \cellcolor{white}}
\newcommand{\RowColorY}{\rowcolor{yellow} \cellcolor{white}}

\newcommand{\RN}[1]{%
  \textup{\uppercase\expandafter{\romannumeral#1}}%
}

\newcommand\Wider[2][3em]{%
\makebox[\linewidth][c]{%
  \begin{minipage}{\dimexpr\textwidth+#1\relax}
  \raggedright#2
  \end{minipage}%
  }%
}

\title{Drought and its Effects on Economy}  
\author{Monika Novackova} 
\date{December 2017} 
\institute{University of Sussex}

\setbeamercolor{block title}{bg=purple!30,fg=black}

\begin{document}

\definecolor{LightCyan}{rgb}{0.88,1,1}
\definecolor{BeigeNadine}{rgb}{1,0.94,0.86}
\definecolor{DarkGreen}{rgb}{0.1803922,0.545098,0.3411765}
\definecolor{BeigeNadine}{rgb}{1,0.94,0.86}
\definecolor{PaleGreen}{rgb}{ 0.5960784,0.9843137, 0.5960784}
\definecolor{darkblue}{rgb}{0.0, 0.0, 0.55}
\definecolor{bored}{rgb}{0.8, 0.0, 0.0}
\definecolor{babypink}{rgb}{0.96, 0.76, 0.76}

 \newcolumntype{d}{D{.}{.}{-1}}
 \newcolumntype{e}{D{+}{\,\pm\,}{6,2}}

\maketitle

\normalsize

\begin{frame}\label{My Interest}

\large{My research interest:} \\
\vspace{0.7cm}

\large{\hspace{1cm}\textbf{Effects of droughts on economy in Kenya}}
\end{frame}
\note{A}

\begin{frame}
\frametitle{Outline}
\tableofcontents
\end{frame}

\note{A}

\section{Definition of Drought}

\begin{frame}\label{Definition of Drought}
\frametitle{Definition of Drought} 
 \begin{itemize}
\item Hi
\end{itemize}

\end{frame}

\note{a\\
b}

\begin{frame}

\frametitle{Effects of droughts on economy}\label{Effects} 
\begin{block}{\textbf{Computable General Equilibrium (CGE)}}

\begin{itemize}

\item \underline{\textbf{\cite{robinson2010}}}
\begin{itemize}
\item 5 agro-ecological zones,46 production activities (incl. 35 zone specific agricultural production sectors), 22 commodity groups,  15 primary factors of production
\end{itemize}

\begin{table}
\begin{footnotesize}
\begin{tabular}{lll}

\hline
\rowcolor{PaleGreen} \textbf{Fixed (inputs)} & \textbf{Determined by model (outputs)}  \\
 \hline
Capital stock& Domestic price of each commodity \\
Land (by region) & Land allocated across crops \\
Supply of labor per skill type& Real wages\\
Foreign capital inflow	& Real exchange rate\\
Trade balance	& \\
 \hline
\end{tabular}
\end{footnotesize}
\end{table}

\begin{footnotesize}

\item The simulation use a 'balanced' macro closure in which aggregate \textbf{investment, government demand, and consumption are fixed shares of total absorption}
\item Intermediate inputs into production are determined as fixed shares
of the quantity of output

\end{footnotesize}
\end{itemize}
\end{block}
\end{frame}
\note{

\vspace{0.3cm}
\vspace{0.3cm}a
\\
\vspace{0.3cm}a
\\
\vspace{0.3cm}a
}

\begin{frame}

\frametitle{Effects of droughts on economy}\label{Effects} 
\begin{block}{\textbf{Computable General Equilibrium (CGE)} Models }
\begin{itemize}

\item \underline{\textbf{\cite{OxfamIDS}}}
\begin{itemize}
\item Exploring range of scenarios for food price increase in 2030
\begin{itemize}
\item 1. Baseline 2. Climate change 3. Climate change with adaptation 4. Adaptation only in sub-Saharan Africa
\end{itemize}
\item Global coverage, set of individual country models, linked through international trade
\item Climate change (incl. drought) modelled as changes in factor productivity (usually negative)

\end{itemize}

\end{itemize}
\end{block}
\end{frame}

\note{A}

\section{My Suggestion}

\begin{frame}

\frametitle{My suggestion - panel estimation}\label{MySuggestion1}

My interest: \textbf{Effects of drought on economy in Kenya}
\begin{itemize}

\item \textbf{Response variable}

\begin{itemize}
\item Volumes of production (crop specific, total)
\item Profit per acre \normalsize{\citep{Deschenes2007Ric}}

\begin{itemize}
\item \footnotesize{(Value of agricultural products - prod. expanses)/acres (crops, pasture, grazing)}\\ 
\end{itemize}
\end{itemize}
\item \textbf{Units of analysis}
\begin{itemize}
\item Counties in Kenya $\times$ year
\
\end{itemize}

\item \textbf{Explanatory variable of interest}

\begin{itemize}
\item Dummy variable (0/1) drought occurred in a particular county and year or not
\item Several varieties - various specifications of drought:

\end{itemize}

\end{itemize}

\end{frame}

\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county\\
\vspace{0.3cm}a\\
\vspace{0.3cm}}

\begin{frame}

\frametitle{My suggestion - panel estimation}\label{MySuggestion3}

My interest: \textbf{Effects of drought on economy in Kenya}
\begin{itemize}

\item \textbf{Explanatory variable of interest}

\begin{itemize}
\item Dummy variable (0/1) drought occurred in a particular county and year or not

\end{itemize}

\item \textbf{Units of analysis:} Counties in Kenya $\times$ year

\item \textbf{Control variables ? - subject to availability}

\begin{itemize}
\item GDP, soil quality data, population,average land area, average value assets,  climate, degree days?
\end{itemize}

\item \textbf{Estimation methods}

\begin{itemize}
\item Fixed effects, SURE, Kalman filter, Box-Jenkins
\end{itemize}

\end{itemize}

\end{frame}

\note{also I can try some t-tests, f-test for various grouping - WHAT WOULD BE THE BEST DEFFINITION OF DROUGHT year x county\\
\vspace{0.3cm}a\\
\vspace{0.3cm}}

\begin{frame}

\frametitle{My idea - panel estimation}\label{MySuggestion4}

\begin{align*}
\begin{array}{lcl}
 Y_{i,t} &=& \alpha_i +\gamma_t + \delta D'_{i,t} +  \boldsymbol{\beta X_{i,t}}+ \epsilon_{i,t}
\end{array}
\end{align*}

\begin{table}
\begin{small}
\begin{tabular}{lll}

$Y_{i,t}$ &$=$& Response variable (food production/price), county $i$ in year $t$\\
$\alpha_i$ &$=$& Fixed effects, county $i$\\
{\color{darkblue}$\delta$ }&{\color{darkblue}$=$}&{\color{darkblue}{Effect of drought on economy}}\\
$D_{i,t}$ &$=$& Indicator variable\\
& &  $D=1$ if drought in county $i$ in year $t$, $D=0$ otherwise\\
$\boldsymbol{\beta}$ &$=$& Vector of effects of other covariates\\
$X_{i,t}$ &$=$& Matrix of values of other covariates in county $i$ in year $t$\\
$\epsilon_{i,t}$ &$=$& Error term \\
{\color{DarkGreen}$\gamma_t$} & {\color{DarkGreen}$=$}& {\color{DarkGreen}Year specific indicator?}
\end{tabular}
\end{small}
\end{table}

\end{frame}

\note{Also waldtest, jako ekvivalent pro celkovy f-test byl signifikantni>> overall fit je moc o}

\note{hello}

\section*{}

\frame[allowframebreaks]{ 
\centering
\huge{Thank you for attention}
}

\note{hello}

\frame[allowframebreaks]{ 
\small
\frametitle{References}
\bibliographystyle{apa}
\bibliography{referencesFS}
}

\end{document}